Reference Tracking Strategies of Deaf Adult Signers in Turkish Sign Language

### 2023-05-04 revised_manuscript.Rmd
updated

In our first set of hypotheses regarding referent tracking strategies, we had the general expectation that signers would use more explicit forms (mainly nominals) for referent introduction and less implicit forms such as zero anaphora, pronominals, and classifiers for maintenance. In re-introduced contexts, we predicted that signers would use both explicit and implicit markers depending on the accessibility of the tracked referent.

For age of acquisition effects, based on what is proposed for child and adult deaf signers with delayed language exposure in terms of their morphosyntactic [@boudreault2006; @cormier2012; @kayaba2022] and narrative development [@becker2009; @cormier2013], we expected that DoH signers would be more over-redundant in their referent tracking, perhaps akin to second language learners [@frederiksen2019; @bel2015]. That is, we had the expectation that DoH signers would use more explicit markers like nouns and fewer implicit markers like zero anaphora even for highly accessible referents (when the discourse status is maintained or even re-introduced in same cases) than DoD signers.

@@

The results of the present study only partly support our hypotheses. In line with our expectations, regardless of the participant groups (DoD or DoH) (i) signers mainly used nominals to introduce a referent into the discourse, (ii) used zero anaphora and classifiers but not pronouns for maintained contexts, and (iii) employed both nominals and less explicit markers (classifiers, zero anaphora and pronouns) to re-introduce a referent.

However, the proportion of the observed discourse statuses (26% for introduction, 40% for maintenance, 34% for re-introduction across acquisition groups) in this study differed from that of some other studies. For a comparison, while the percentage of introduced referents were similar in both research (26% versus 24%), the participants in the present study brought back referents into the discourse more frequently compared to the participants in @frederiksen2019 and @frederiksen2016 which reported that re-introduction contexts accounted for only 7% of the referents in ASL. On the other hand, maintenance percentages were lower in the present study compared to @frederiksen2016 (40% versus 69%). Investigating the distribution of REs across discourse statuses in ASL, a similar study by @czubek2017 also reported contrasting numbers (17% for introduction, 64% for maintenance, and 19% for re-introduction). We suggest that the discrepancy might be due to the nature of the stimuli. In @frederiksen2016, the authors used a 4-6-episode-long picture story with 2 animate characters and 1 inanimate object. In contrast, the stimuli in the present study consisted of 8-episode-long video clips on average with a mean number of 2 animate characters and 3 inanimate objects. Although the number of animate characters and inanimate characters in Czubek's [-@czubek2017] stimuli matched that of the present study, their narrative structure (a 6-episode-long picture story) was still shorter compared to our narrative stimuli. We believe fewer possible referents and/or a simpler narrative structure in these studies may have contributed to the lower number of re-introduced contexts compared to the present study. Due to the story length and composition in this study, the participants had to track more recurring animate or inanimate referents for a longer period of time. As a result, our participants might have engaged in referent re-introduction more than what was reported in previous studies [@czubek2017; @frederiksen2016].

@@

Interestingly and to our surprise, the signers in this study used pronominal IX very sporadically with only 40 observed instances that were mainly used to re-introduce referents rather than maintain them. This finding is comparable with what has been reported for ASL [@swabey2002; @frederiksen2016; @czubek2017] but not with LSC [@bel2015], BSL [@morgan2005] and DGS [@perniss2015], in which the signers extensively made overt pronominal references (up to 30%) to maintain entities. Comparable to our findings, it has been reported that pronouns do not frequently occur in the local context and instead follows a topic shift and thus rank higher than classifiers and zero anaphora [see @barbera2009 for LSC; and @nuhbalaoglu2018 for TİD]. Considering this, the claim that the IX sign functions like demonstratives that are also used to track referents in re-introduced contexts has not been borne out [@koulidobrova2016]. Similar to these research, our findings indicate that the pronominal IX in TİD was used more concurrently with nominal constructions and pronominal IX was used very infrequently in isolation. The observation that overt pronouns are rare in ASL narratives comes from studies that include a simple picture retelling task with one protagonist [e.g., @frederiksen2016] as well as a more complex narrative design as in [@czubek2017]. However, @czubek2017 also incorporated different genres and reported more instances of overt pronouns in personal narratives (compared to picture or video retelling). In fact, this finding suggests that the use of pronouns might be genre-dependent. Moreover, another production study on referent tracking in ASL reported object preference for pronominals [@frederiksen2017]. However, we only coded subjects for maintained and re-introduced referents. We argue that all of these factors (genre-dependency as well as object preference) might contribute to the explanation of why overt pronouns in our data are dispreferred in favor of other markers for highly accessible referents such as zero anaphora and classifiers.

A deeper look into the allocation of classifier handshapes revealed that WECLs were the most common type of classifiers for the maintenance of a referent followed by BPCLs and ExtCLs. We observed a similar distibution for re-introduced contexts. In contrast, a great number of the overall ExtCLs were used to introduce a new referent into the discourse. That is to say, whereas WECLs were the preferred classifier types to signal highly accessible referents, ExtCL constructions were mainly used for referents which were not easily accessible. We suggest that the less iconic and transparent nature of WECLs allowed signers to continue a previous referent without having to use an explict form. As pointed out by @czubek2017, signers prefer to rely on ExtCLs for the first mention of an entity because these constructions are informative and explicit by nature, and they provide visual and descriptive cues about an entity and its location in space. We only observed `r sum(df_narration$GTR1_plot == "SASS" & df_narration$Discourse == "Re-introduction")` instances of ExtCLs for re-introduced contexts, replicating Frederiksen and Mayberry's [-@frederiksen2019] findings for ASL.

@@

% Text Node
\draw (498,100) node [anchor=north west][inner sep=0.75pt]   [align=left] {\textbf{High}};
% Text Node
\draw (81,100) node [anchor=north west][inner sep=0.75pt]   [align=left] {\textbf{Low}};
% Text Node
\draw (150,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize NOM}};
% Text Node
\draw (250,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize ExtCL}};
% Text Node% Text Node
\draw (333,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize AGR \& CA}};
% Text Node
\draw (445,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize PLAIN}};
\end{tikzpicture}
\begin{minipage}{\textwidth}
    \footnotesize
    \centering
    Accessibility hierarchy in Frederiksen and Mayberry (2016) for ASL
    \end{minipage}
\end{figure}
```
Both hierarchies illustrated in Figure \ref{fig-scale} treat nominals and ExtCLs as markers of lowly accessibly referents. Likewise, in both hierarchies, as markers of highly accessible referents, different types of zero anaphora markers are further distinguished. Plain verbs rank the lowest on the scales, preceded by agreement verbs and CA. On a similar ground, both scales lack pronouns, which occurred very infrequently in the data. Although @frederiksen2016 did not include other classifier types in the hierarchy, we propose that WECLs and BPCLs must rank higher than zero anaphora but lower than ExtCLs that were mainly used by interlocutors to introduce a referent into the discourse context.

@@

The results in this study only somewhat support our second set of hypotheses that we set for the effects of having signing deaf parents on participants' RE selection. That is, while being a DoD signer of TİD increased the odds of using zero anaphora in reference tracking across different discourse statuses, we failed to find a difference for overall nominal production. Taking a closer look at the interaction terms, we observed that DoD signers, compared to DoH signers, were less likely to use nominal constructions when the discourse status was maintained. Native acquisition from deaf caregivers further increased the odds of zero anaphora when the discourse status was introduced. However, we found the opposite alignment for the interaction of maintenance and native acquisition. That is, the coefficient for this interaction term slightly decreased the combined main effects of having a maintained context and being a DoD signer. This means that these main effects were greater than the interaction or added effects, pointing out to the implication that the type of discourse status (introduced, maintained, re-introduced) accounts for the distribution of REs in TİD narratives better than native acquisition from deaf caregivers (DoD, DoH) which, in comparison, had a less salient effect.

While DoH signers were only partially over-redundant in maintained contexts, they were more so in introduced contexts. In line with our study, a tendency for limited over-redundancy has been suggested for late signing children in DGS [@becker2009] and late signing adults in BSL specifically for the use of CA [@cormier2013]. As for TİD, to our knowledge, there is only one study [@gur2018] that directly aimed at examining age of acquisition effects on adult signers' narrative skills. @gur2018 observed that late signing deaf adults and children set the scene less frequently and used CA less accurately compared to their native counterparts. However, this particular investigation did not present any data on the distribution of REs between two acquisition groups. In the present study, the finding that DoD signers preferred to use null arguments even for the first mention of a referent (`r sum(df_narration$ref_type == "Zero" & df_narration$Discourse == "Introduction")` in total) is surprising given some of the previous literature in the topicality and saliency framework [@cormier2013; @swabey2002; @ferrara2022]. These research reported very little or no use of zero anaphora by DoD signers for introduced contexts.

>>>>

Regarding referent tracking strategies, we had the general expectation that signers would use more explicit forms (mainly nominals) for referent introduction and less implicit forms such as zero anaphora, pronominals, and classifiers for maintenance. In re-introduced contexts, we predicted that signers would use both explicit and implicit markers depending on the accessibility of the tracked referent.

Especially, for age of acquisition effects, based on what is proposed for child and adult deaf signers with delayed language exposure in terms of their morphosyntactic [@boudreault2006; @cormier2012; @kayaba2022] and narrative development [@becker2009; @cormier2013], we expected that DoH signers would be more over-redundant in their referent tracking, perhaps akin to second language learners [@frederiksen2019; @bel2015]. That is, we had the expectation that DoH signers would use more explicit markers like nouns and fewer implicit markers like zero anaphora even for highly accessible referents (when the discourse status is maintained or even re-introduced in same cases) than DoD signers.

@@

Mainly, the results of the present study support our initial set of hypotheses. In line with our expectations, regardless of the participant groups (DoD or DoH) (i) signers mainly used nominals to introduce a referent into the discourse, (ii) used zero anaphora and classifiers but not pronouns for maintained contexts, and (iii) employed both nominals and less explicit markers (classifiers, zero anaphora and pronouns) to re-introduce a referent.

By contrast, the proportion of the observed discourse statuses (26% for introduction, 40% for maintenance, 34% for re-introduction across acquisition groups) in this study differed from that of some other studies. For a comparison, while the percentage of introduced referents were similar in both research (26% versus 24%), the participants in the present study brought back referents into the discourse more frequently compared to the participants in @frederiksen2019 and @frederiksen2016 which reported that re-introduction contexts accounted for only 7% of the referents in ASL. On the other hand, maintenance percentages were lower in the present study compared to @frederiksen2016 (40% versus 69%). Investigating the distribution of REs across discourse statuses in ASL, a similar study by @czubek2017 also reported contrasting numbers (17% for introduction, 64% for maintenance, and 19% for re-introduction). We suggest that the discrepancy might be due to the nature of the stimuli. In @frederiksen2016, the authors used a 4-6-episode-long picture story with 2 animate characters and 1 inanimate object. In contrast, the stimuli in the present study consisted of 8-episode-long video clips on average with a mean number of 2 animate characters and 3 inanimate objects. Although the number of animate characters and inanimate characters in Czubek's [-@czubek2017] stimuli matched that of the present study, their narrative structure (a 6-episode-long picture story) was still shorter compared to our narrative stimuli. We believe fewer possible referents and/or a simpler narrative structure in these studies may have contributed to the lower number of re-introduced contexts compared to the present study. Due to the story length and composition in this study, the participants had to track more recurring animate or inanimate referents for a longer period of time. As a result, our participants might have engaged in referent re-introduction more than what was reported in previous studies [@czubek2017; @frederiksen2016].

@@

Yet, surprisingly, the signers in this study used pronominal IX very sporadically with only 40 observed instances that were mainly used to re-introduce referents rather than maintain them. This finding is comparable with what has been reported for ASL [@swabey2002; @frederiksen2016; @czubek2017] but not with LSC [@bel2015], BSL [@morgan2005] and DGS [@perniss2015], in which the signers extensively made overt pronominal references (up to 30%) to maintain entities. Comparable to our findings, it has been reported that pronouns do not frequently occur in the local context and instead follows a topic shift and thus rank higher than classifiers and zero anaphora [see @barbera2009 for LSC; and @nuhbalaoglu2018 for TİD]. Considering this, the claim that the IX sign functions like demonstratives that are also used to track referents in re-introduced contexts has not been borne out [@koulidobrova2016]. Similar to these research, our findings indicate that the pronominal IX in TİD was used more concurrently with nominal constructions and pronominal IX was used very infrequently in isolation. The observation that overt pronouns are rare in ASL narratives comes from studies that include a simple picture retelling task with one protagonist [e.g., @frederiksen2016] as well as a more complex narrative design as in [@czubek2017]. However, @czubek2017 also incorporated different genres and reported more instances of overt pronouns in personal narratives (compared to picture or video retelling). In fact, this finding suggests that the use of pronouns might be genre-dependent. Moreover, another production study on referent tracking in ASL reported object preference for pronominals [@frederiksen2017]. However, we only coded subjects for maintained and re-introduced referents. We argue that all of these factors (genre-dependency as well as object preference) might contribute to the explanation of why overt pronouns in our data are dispreferred in favor of other markers for highly accessible referents such as zero anaphora and classifiers.

Regarding the allocation of classifier handshapes, we found that WECLs were the most common type of classifiers for the maintenance of a referent followed by BPCLs and ExtCLs. We observed a similar distibution for re-introduced contexts. In contrast, a great number of the overall ExtCLs were used to introduce a new referent into the discourse. That is to say, whereas WECLs were the preferred classifier types to signal highly accessible referents, ExtCL constructions were mainly used for referents which were not easily accessible. We suggest that the less iconic and transparent nature of WECLs allowed signers to continue a previous referent without having to use an explict form. As pointed out by @czubek2017, signers prefer to rely on ExtCLs for the first mention of an entity because these constructions are informative and explicit by nature, and they provide visual and descriptive cues about an entity and its location in space. We only observed `r sum(df_narration$GTR1_plot == "SASS" & df_narration$Discourse == "Re-introduction")` instances of ExtCLs for re-introduced contexts, replicating Frederiksen and Mayberry's [-@frederiksen2019] findings for ASL.

@@

% Text Node
\draw (498,100) node [anchor=north west][inner sep=0.75pt]   [align=left] {\textbf{High}};
% Text Node
\draw (81,100) node [anchor=north west][inner sep=0.75pt]   [align=left] {\textbf{Low}};
% Text Node
\draw (150,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize NOM}};
% Text Node
\draw (250,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize ExtCL}};
% Text Node% Text Node
\draw (333,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize AGR \& CA}};
% Text Node
\draw (445,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize PLAIN}};
\end{tikzpicture}
\begin{minipage}{\textwidth}
    \footnotesize
    \centering
    Accessibility hierarchy in Frederiksen and Mayberry (2016) for ASL
    \end{minipage}
\end{figure}
```
Keep in mind that both hierarchies illustrated in Figure \ref{fig-scale} treat nominals and ExtCLs as markers of lowly accessibly referents. In both hierarchies, as markers of highly accessible referents, different types of zero anaphora markers are further distinguished. Plain verbs rank the lowest on the scales, preceded by agreement verbs and CA. On a similar ground, both scales lack pronouns, which occurred very infrequently in the data. Although @frederiksen2016 did not include other classifier types in the hierarchy, we propose that WECLs and BPCLs must rank higher than zero anaphora but lower than ExtCLs that were mainly used by interlocutors to introduce a referent into the discourse context.

@@

Likewise, our results only partly support the second set of hypotheses that we set for the effects of having signing deaf parents on participants' RE selection. That is, while being a DoD signer of TİD increased the odds of using zero anaphora in reference tracking across different discourse statuses, we failed to find a difference for overall nominal production. Taking a closer look at the interaction terms, we observed that DoD signers, compared to DoH signers, were less likely to use nominal constructions when the discourse status was maintained. Native acquisition from deaf caregivers further increased the odds of zero anaphora when the discourse status was introduced. However, we found the opposite alignment for the interaction of maintenance and native acquisition. That is, the coefficient for this interaction term slightly decreased the combined main effects of having a maintained context and being a DoD signer. This means that these main effects were greater than the interaction or added effects, pointing out to the implication that the type of discourse status (introduced, maintained, re-introduced) accounts for the distribution of REs in TİD narratives better than native acquisition from deaf caregivers (DoD, DoH) which, in comparison, had a less salient effect.

It can be observed that while DoH signers were only partially over-redundant in maintained contexts, they were more so in introduced contexts. In line with our study, a tendency for limited over-redundancy has been suggested for late signing children in DGS [@becker2009] and late signing adults in BSL specifically for the use of CA [@cormier2013]. As for TİD, to our knowledge, there is only one study [@gur2018] that directly aimed at examining age of acquisition effects on adult signers' narrative skills. @gur2018 observed that late signing deaf adults and children set the scene less frequently and used CA less accurately compared to their native counterparts. However, this particular investigation did not present any data on the distribution of REs between two acquisition groups. In the present study, the finding that DoD signers preferred to use null arguments even for the first mention of a referent (`r sum(df_narration$ref_type == "Zero" & df_narration$Discourse == "Introduction")` in total) is surprising given some of the previous literature in the topicality and saliency framework [@cormier2013; @swabey2002; @ferrara2022]. These research reported very little or no use of zero anaphora by DoD signers for introduced contexts.

### 2023-05-04 revised_manuscript.tex
updated

In our first set of hypotheses regarding referent tracking strategies,
we had the general expectation that signers would use more explicit
forms (mainly nominals) for referent introduction and less implicit
forms such as zero anaphora, pronominals, and classifiers for
maintenance. In re-introduced contexts, we predicted that signers would
use both explicit and implicit markers depending on the accessibility of
the tracked referent.

For age of acquisition effects, based on what is proposed for child and
adult deaf signers with delayed language exposure in terms of their
morphosyntactic (Boudreault and Mayberry, 2006; Cormier et al., 2012;
Kayabaşı and Gökgöz, 2022) and narrative development (Becker, 2009;
Cormier et al., 2013), we expected that DoH signers would be more
over-redundant in their referent tracking, perhaps akin to second
language learners (Bel et al., 2015; Frederiksen and Mayberry, 2019).
That is, we had the expectation that DoH signers would use more explicit
markers like nouns and fewer implicit markers like zero anaphora even
for highly accessible referents (when the discourse status is maintained
or even re-introduced in same cases) than DoD signers.

@@

The results of the present study only partly support our hypotheses. In
line with our expectations, regardless of the participant groups (DoD or
DoH) (i) signers mainly used nominals to introduce a referent into the
discourse, (ii) used zero anaphora and classifiers but not pronouns for
maintained contexts, and (iii) employed both nominals and less explicit
markers (classifiers, zero anaphora and pronouns) to re-introduce a
referent.

However, the proportion of the observed discourse statuses (26\% for
introduction, 40\% for maintenance, 34\% for re-introduction across
acquisition groups) in this study differed from that of some other
studies. For a comparison, while the percentage of introduced referents
were similar in both research (26\% versus 24\%), the participants in
the present study brought back referents into the discourse more
frequently compared to the participants in Frederiksen and Mayberry
(2019) and Frederiksen and Mayberry (2016) which reported that
re-introduction contexts accounted for only 7\% of the referents in ASL.
On the other hand, maintenance percentages were lower in the present
study compared to Frederiksen and Mayberry (2016) (40\% versus 69\%).
Investigating the distribution of REs across discourse statuses in ASL,
a similar study by Czubek (2017) also reported contrasting numbers (17\%
for introduction, 64\% for maintenance, and 19\% for re-introduction).
We suggest that the discrepancy might be due to the nature of the
stimuli. In Frederiksen and Mayberry (2016), the authors used a
4-6-episode-long picture story with 2 animate characters and 1 inanimate
object. In contrast, the stimuli in the present study consisted of
8-episode-long video clips on average with a mean number of 2 animate
characters and 3 inanimate objects. Although the number of animate
characters and inanimate characters in Czubek's (2017) stimuli matched
that of the present study, their narrative structure (a 6-episode-long
picture story) was still shorter compared to our narrative stimuli. We
believe fewer possible referents and/or a simpler narrative structure in
these studies may have contributed to the lower number of re-introduced
contexts compared to the present study. Due to the story length and
composition in this study, the participants had to track more recurring
animate or inanimate referents for a longer period of time. As a result,
our participants might have engaged in referent re-introduction more
than what was reported in previous studies (Czubek, 2017; Frederiksen
and Mayberry, 2016).

@@

Interestingly and to our surprise, the signers in this study used
pronominal IX very sporadically with only 40 observed instances that
were mainly used to re-introduce referents rather than maintain them.
This finding is comparable with what has been reported for ASL (Czubek,
2017; Frederiksen and Mayberry, 2016; Swabey, 2002) but not with LSC
(Bel et al., 2015), BSL (Morgan, 2005) and DGS (Perniss and Özyürek,
2015), in which the signers extensively made overt pronominal references
(up to 30\%) to maintain entities. Comparable to our findings, it has
been reported that pronouns do not frequently occur in the local context
and instead follows a topic shift and thus rank higher than classifiers
and zero anaphora (see Barberà and Massó, 2009 for LSC; and Nuhbalaoglu,
2018 for TİD). Considering this, the claim that the IX sign functions
like demonstratives that are also used to track referents in
re-introduced contexts has not been borne out (Koulidobrova and
Lillo-Martin, 2016). Similar to these research, our findings indicate
that the pronominal IX in TİD was used more concurrently with nominal
constructions and pronominal IX was used very infrequently in isolation.
The observation that overt pronouns are rare in ASL narratives comes
from studies that include a simple picture retelling task with one
protagonist (e.g., Frederiksen and Mayberry, 2016) as well as a more
complex narrative design as in (Czubek, 2017). However, Czubek (2017)
also incorporated different genres and reported more instances of overt
pronouns in personal narratives (compared to picture or video
retelling). In fact, this finding suggests that the use of pronouns
might be genre-dependent. Moreover, another production study on referent
tracking in ASL reported object preference for pronominals (Frederiksen
and Mayberry, 2017). However, we only coded subjects for maintained and
re-introduced referents. We argue that all of these factors
(genre-dependency as well as object preference) might contribute to the
explanation of why overt pronouns in our data are dispreferred in favor
of other markers for highly accessible referents such as zero anaphora
and classifiers.

A deeper look into the allocation of classifier handshapes revealed that
WECLs were the most common type of classifiers for the maintenance of a
referent followed by BPCLs and ExtCLs. We observed a similar distibution
for re-introduced contexts. In contrast, a great number of the overall
ExtCLs were used to introduce a new referent into the discourse. That is
to say, whereas WECLs were the preferred classifier types to signal
highly accessible referents, ExtCL constructions were mainly used for
referents which were not easily accessible. We suggest that the less
iconic and transparent nature of WECLs allowed signers to continue a
previous referent without having to use an explict form. As pointed out
by Czubek (2017), signers prefer to rely on ExtCLs for the first mention
of an entity because these constructions are informative and explicit by
nature, and they provide visual and descriptive cues about an entity and
its location in space. We only observed 3 instances of ExtCLs for
re-introduced contexts, replicating Frederiksen and Mayberry's (2019)
findings for ASL.

@@

Both hierarchies illustrated in Figure \ref{fig-scale} treat nominals
and ExtCLs as markers of lowly accessibly referents. Likewise, in both
hierarchies, as markers of highly accessible referents, different types
of zero anaphora markers are further distinguished. Plain verbs rank the
lowest on the scales, preceded by agreement verbs and CA. On a similar
ground, both scales lack pronouns, which occurred very infrequently in
the data. Although Frederiksen and Mayberry (2016) did not include other
classifier types in the hierarchy, we propose that WECLs and BPCLs must
rank higher than zero anaphora but lower than ExtCLs that were mainly
used by interlocutors to introduce a referent into the discourse
context.

@@

The results in this study only somewhat support our second set of
hypotheses that we set for the effects of having signing deaf parents on
participants' RE selection. That is, while being a DoD signer of TİD
increased the odds of using zero anaphora in reference tracking across
different discourse statuses, we failed to find a difference for overall
nominal production. Taking a closer look at the interaction terms, we
observed that DoD signers, compared to DoH signers, were less likely to
use nominal constructions when the discourse status was maintained.
Native acquisition from deaf caregivers further increased the odds of
zero anaphora when the discourse status was introduced. However, we
found the opposite alignment for the interaction of maintenance and
native acquisition. That is, the coefficient for this interaction term
slightly decreased the combined main effects of having a maintained
context and being a DoD signer. This means that these main effects were
greater than the interaction or added effects, pointing out to the
implication that the type of discourse status (introduced, maintained,
re-introduced) accounts for the distribution of REs in TİD narratives
better than native acquisition from deaf caregivers (DoD, DoH) which, in
comparison, had a less salient effect.

While DoH signers were only partially over-redundant in maintained
contexts, they were more so in introduced contexts. In line with our
study, a tendency for limited over-redundancy has been suggested for
late signing children in DGS (Becker, 2009) and late signing adults in
BSL specifically for the use of CA (Cormier et al., 2013). As for TİD,
to our knowledge, there is only one study (Gür, 2018) that directly
aimed at examining age of acquisition effects on adult signers'
narrative skills. Gür (2018) observed that late signing deaf adults and
children set the scene less frequently and used CA less accurately
compared to their native counterparts. However, this particular
investigation did not present any data on the distribution of REs
between two acquisition groups. In the present study, the finding that
DoD signers preferred to use null arguments even for the first mention
of a referent (19 in total) is surprising given some of the previous
literature in the topicality and saliency framework (Cormier et al.,
2013; Ferrara et al., 2022; Swabey, 2002). These research reported very
little or no use of zero anaphora by DoD signers for introduced
contexts.

>>>>

Regarding referent tracking strategies, we had the general expectation
that signers would use more explicit forms (mainly nominals) for
referent introduction and less implicit forms such as zero anaphora,
pronominals, and classifiers for maintenance. In re-introduced contexts,
we predicted that signers would use both explicit and implicit markers
depending on the accessibility of the tracked referent.

Especially, for age of acquisition effects, based on what is proposed
for child and adult deaf signers with delayed language exposure in terms
of their morphosyntactic (Boudreault and Mayberry, 2006; Cormier et al.,
2012; Kayabaşı and Gökgöz, 2022) and narrative development (Becker,
2009; Cormier et al., 2013), we expected that DoH signers would be more
over-redundant in their referent tracking, perhaps akin to second
language learners (Bel et al., 2015; Frederiksen and Mayberry, 2019).
That is, we had the expectation that DoH signers would use more explicit
markers like nouns and fewer implicit markers like zero anaphora even
for highly accessible referents (when the discourse status is maintained
or even re-introduced in same cases) than DoD signers.

@@

Mainly, the results of the present study support our initial set of
hypotheses. In line with our expectations, regardless of the participant
groups (DoD or DoH) (i) signers mainly used nominals to introduce a
referent into the discourse, (ii) used zero anaphora and classifiers but
not pronouns for maintained contexts, and (iii) employed both nominals
and less explicit markers (classifiers, zero anaphora and pronouns) to
re-introduce a referent.

By contrast, the proportion of the observed discourse statuses (26\% for
introduction, 40\% for maintenance, 34\% for re-introduction across
acquisition groups) in this study differed from that of some other
studies. For a comparison, while the percentage of introduced referents
were similar in both research (26\% versus 24\%), the participants in
the present study brought back referents into the discourse more
frequently compared to the participants in Frederiksen and Mayberry
(2019) and Frederiksen and Mayberry (2016) which reported that
re-introduction contexts accounted for only 7\% of the referents in ASL.
On the other hand, maintenance percentages were lower in the present
study compared to Frederiksen and Mayberry (2016) (40\% versus 69\%).
Investigating the distribution of REs across discourse statuses in ASL,
a similar study by Czubek (2017) also reported contrasting numbers (17\%
for introduction, 64\% for maintenance, and 19\% for re-introduction).
We suggest that the discrepancy might be due to the nature of the
stimuli. In Frederiksen and Mayberry (2016), the authors used a
4-6-episode-long picture story with 2 animate characters and 1 inanimate
object. In contrast, the stimuli in the present study consisted of
8-episode-long video clips on average with a mean number of 2 animate
characters and 3 inanimate objects. Although the number of animate
characters and inanimate characters in Czubek's (2017) stimuli matched
that of the present study, their narrative structure (a 6-episode-long
picture story) was still shorter compared to our narrative stimuli. We
believe fewer possible referents and/or a simpler narrative structure in
these studies may have contributed to the lower number of re-introduced
contexts compared to the present study. Due to the story length and
composition in this study, the participants had to track more recurring
animate or inanimate referents for a longer period of time. As a result,
our participants might have engaged in referent re-introduction more
than what was reported in previous studies (Czubek, 2017; Frederiksen
and Mayberry, 2016).

@@

Yet, surprisingly, the signers in this study used pronominal IX very
sporadically with only 40 observed instances that were mainly used to
re-introduce referents rather than maintain them. This finding is
comparable with what has been reported for ASL (Czubek, 2017;
Frederiksen and Mayberry, 2016; Swabey, 2002) but not with LSC (Bel et
al., 2015), BSL (Morgan, 2005) and DGS (Perniss and Özyürek, 2015), in
which the signers extensively made overt pronominal references (up to
30\%) to maintain entities. Comparable to our findings, it has been
reported that pronouns do not frequently occur in the local context and
instead follows a topic shift and thus rank higher than classifiers and
zero anaphora (see Barberà and Massó, 2009 for LSC; and Nuhbalaoglu,
2018 for TİD). Considering this, the claim that the IX sign functions
like demonstratives that are also used to track referents in
re-introduced contexts has not been borne out (Koulidobrova and
Lillo-Martin, 2016). Similar to these research, our findings indicate
that the pronominal IX in TİD was used more concurrently with nominal
constructions and pronominal IX was used very infrequently in isolation.
The observation that overt pronouns are rare in ASL narratives comes
from studies that include a simple picture retelling task with one
protagonist (e.g., Frederiksen and Mayberry, 2016) as well as a more
complex narrative design as in (Czubek, 2017). However, Czubek (2017)
also incorporated different genres and reported more instances of overt
pronouns in personal narratives (compared to picture or video
retelling). In fact, this finding suggests that the use of pronouns
might be genre-dependent. Moreover, another production study on referent
tracking in ASL reported object preference for pronominals (Frederiksen
and Mayberry, 2017). However, we only coded subjects for maintained and
re-introduced referents. We argue that all of these factors
(genre-dependency as well as object preference) might contribute to the
explanation of why overt pronouns in our data are dispreferred in favor
of other markers for highly accessible referents such as zero anaphora
and classifiers.

Regarding the allocation of classifier handshapes, we found that WECLs
were the most common type of classifiers for the maintenance of a
referent followed by BPCLs and ExtCLs. We observed a similar distibution
for re-introduced contexts. In contrast, a great number of the overall
ExtCLs were used to introduce a new referent into the discourse. That is
to say, whereas WECLs were the preferred classifier types to signal
highly accessible referents, ExtCL constructions were mainly used for
referents which were not easily accessible. We suggest that the less
iconic and transparent nature of WECLs allowed signers to continue a
previous referent without having to use an explict form. As pointed out
by Czubek (2017), signers prefer to rely on ExtCLs for the first mention
of an entity because these constructions are informative and explicit by
nature, and they provide visual and descriptive cues about an entity and
its location in space. We only observed 3 instances of ExtCLs for
re-introduced contexts, replicating Frederiksen and Mayberry's (2019)
findings for ASL.

@@

Keep in mind that both hierarchies illustrated in Figure \ref{fig-scale}
treat nominals and ExtCLs as markers of lowly accessibly referents. In
both hierarchies, as markers of highly accessible referents, different
types of zero anaphora markers are further distinguished. Plain verbs
rank the lowest on the scales, preceded by agreement verbs and CA. On a
similar ground, both scales lack pronouns, which occurred very
infrequently in the data. Although Frederiksen and Mayberry (2016) did
not include other classifier types in the hierarchy, we propose that
WECLs and BPCLs must rank higher than zero anaphora but lower than
ExtCLs that were mainly used by interlocutors to introduce a referent
into the discourse context.

@@

Likewise, our results only partly support the second set of hypotheses
that we set for the effects of having signing deaf parents on
participants' RE selection. That is, while being a DoD signer of TİD
increased the odds of using zero anaphora in reference tracking across
different discourse statuses, we failed to find a difference for overall
nominal production. Taking a closer look at the interaction terms, we
observed that DoD signers, compared to DoH signers, were less likely to
use nominal constructions when the discourse status was maintained.
Native acquisition from deaf caregivers further increased the odds of
zero anaphora when the discourse status was introduced. However, we
found the opposite alignment for the interaction of maintenance and
native acquisition. That is, the coefficient for this interaction term
slightly decreased the combined main effects of having a maintained
context and being a DoD signer. This means that these main effects were
greater than the interaction or added effects, pointing out to the
implication that the type of discourse status (introduced, maintained,
re-introduced) accounts for the distribution of REs in TİD narratives
better than native acquisition from deaf caregivers (DoD, DoH) which, in
comparison, had a less salient effect.

It can be observed that while DoH signers were only partially
over-redundant in maintained contexts, they were more so in introduced
contexts. In line with our study, a tendency for limited over-redundancy
has been suggested for late signing children in DGS (Becker, 2009) and
late signing adults in BSL specifically for the use of CA (Cormier et
al., 2013). As for TİD, to our knowledge, there is only one study (Gür,
2018) that directly aimed at examining age of acquisition effects on
adult signers' narrative skills. Gür (2018) observed that late signing
deaf adults and children set the scene less frequently and used CA less
accurately compared to their native counterparts. However, this
particular investigation did not present any data on the distribution of
REs between two acquisition groups. In the present study, the finding
that DoD signers preferred to use null arguments even for the first
mention of a referent (19 in total) is surprising given some of the
previous literature in the topicality and saliency framework (Cormier et
al., 2013; Ferrara et al., 2022; Swabey, 2002). These research reported
very little or no use of zero anaphora by DoD signers for introduced
contexts.
